---
title: "homework dv 5"
format: html
editor: visual
embed-resources: true
---

## Homework DV 5

```{r}
library(tidyverse)
# Read in the data 
wncaa <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/wncaa.csv")

# Glimpse the data 
glimpse(wncaa)

```

## Question 1

```{r}
wncaa_champ <-  wncaa %>%
  filter(tourney_finish == "Champ")
```

```{r}
percentage_wins <- wncaa_champ %>%
  group_by(school) %>%
  summarize(N = n()) %>%
  mutate(percent = N / sum(N),
         pct = round(percent*100), 0)
```

```{r}
ggplot(data = percentage_wins, mapping = aes(x = school,
                                             y = pct)) +
  geom_col(position = "dodge2", fill = "navy") +
  scale_y_continuous(limits = c(0, 100)) +
  coord_flip() +
  labs(title = "Tourney Wins by %", x = "School", y = "Win %")
```

## Question 2 

```{r}
champ_names <- unique(wncaa_champ$school)
```

```{r}
winners <- wncaa %>% 
  filter(school %in% champ_names)
```

```{r}
ggplot(data = winners, mapping = aes(x = school,
                                     y = seed,
                                     fill = school)) +
  geom_boxplot(outlier.shape = NA) +
  coord_flip() +
  labs(title = "Seeds by School", x = "school", y = "seed", caption = "source: wncaa")
```

```{r}
ggplot(data = winners, mapping = aes(x = school,
                                     y = seed,
                                     fill = school)) +
  geom_jitter(alpha = 0.25) +
  geom_boxplot(outlier.shape = NA) +
  coord_flip() +
  labs(title = "Seeds by School", x = "School", y = "Seed", caption = "Source: WNCAA")
```

```{r}
top_winners <- winners |>   
  group_by(school) %>%  
  mutate(mean_seed = mean(seed)) %>% 
  ungroup()
```

```{r}
ggplot(top_winners, 
       aes(x = reorder(school, -mean_seed), 
           y = seed, 
           fill = school)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 0.25) +
  coord_flip() +
  labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "Source: WNCAA")
```

From this new graph, I can see that higher seeds (closer to 1) have few outliers, which makes sense because these seeds have won the most titles (i.e. UConn, Tennessee)

```{r}
ggplot(top_winners, 
       aes(x = reorder(school, -mean_seed), 
           y = seed, 
           fill = school)) +
  geom_violin() +
  coord_flip() +
  labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "Source: WNCAA")
```

I would say the previous graphs with the box plots are more informative because it clearly establishes a median and the number of outliers of each team. The "geom_violin" graph is a more "summarized" version of the box plot graphs that makes the locations of outliers and medians less clear.

## Question 3

```{r}
ggplot(top_winners, 
       aes(x = reorder(school, -mean_seed), 
           y = seed, 
           fill = school)) +
  geom_point() +
  geom_jitter(alpha = 0.25) +
  coord_flip() +
  labs(title = "Distribution of Seeds by School", x = "School", y = "Seed", caption = "Source: WNCAA")


```

Using geom_point, the above graph doesn't clearly show the medians and outliers of the different seeds by school.

## Question 4

```{r}
winners_mean_sd <- winners |> 
  group_by(school) |> 
  summarize(across(where(is.numeric),
                   list(mean = mean,
                        sd = sd)))
```

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_point() +
  coord_flip() +
  labs(title = "Reg Season Win % by School", x = "School", y = "Avg Reg Season Win %", caption = "source: WNCAA")
```

Analyzing the graph above, I can see that Texas A&M had the lowest average regular season win percentage with a win percentage of about 74%. UConn had the highest with an average regular season win percentage of about 91%.

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_pointrange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd,
                                ymax = reg_percent_mean + reg_percent_sd)) +
  coord_flip() +
  labs(title = "Reg Season Win % by School with SD", x = "School", y = "Avg Reg Season Win %", caption = "source: WNCAA")
```

The school with the narrowest interval is Texas A&M.

```{r}
ggplot(winners_mean_sd,
        mapping = aes(x = reorder(school, reg_percent_mean, na.rm = TRUE),
                      y = reg_percent_mean)) +
  geom_linerange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd,
                               ymax = reg_percent_mean + reg_percent_sd)) +
  coord_flip() +
  labs(title = "Reg Season Win % by School with SD", x = "School", y = "Avg Reg Season Win %", caption = "source: WNCAA")
```

This graph seems similar, but isn't useful because it provide the mean of the regular season average win percentage.
