---
title: "HW MD 6.2"
format: html
editor: visual
embed-resources: true
---

```{r}
library(tidyverse)
library(moderndive)
theme_set(theme_minimal())
```

```{r}
data(bikes, package = "bayesrules")
glimpse(bikes)
```

## Question 0

1.  500 rows in the data-set
2.  Each row represents an observation of a separate data entry
3.  Most days of the year from the years 2011 and 2012 (about 65%)
4.  6946 rides
5.  34.00002

## Question 1

```{r}
bikes %>% 
  select(rides, temp_feel) %>% 
  cor(use = "complete.obs")
```

The correlation between rides and temp_feel is 0.5824898

```{r}
bikes %>% 
  select(rides, windspeed) %>% 
  cor(use = "complete.obs")
```

The correlation between rides and windspeed is -0.1949352

## Question 2

```{r}
bikes <- bikes %>%
  mutate(wind_kph=windspeed*1.61)
```

```{r}
bikes %>% 
  select(windspeed, wind_kph) %>% 
  cor(use = "complete.obs")
```

The correlation between wind-speed and wind-speed in kph is 1. This is because ??????

## Question 3

```{r}
rides_4mph <- lm(rides ~ windspeed, data = bikes)

broom::tidy(rides_4mph)
```

```{r}
rides_4kph <- lm(rides ~ wind_kph, data = bikes)

broom::tidy(rides_4kph)
```

## Question 4

```{r}
pred_df <- tribble(         # tribble allows easy manual data entry
  ~wind_kph,
  20)

pred_df <- pred_df |>
  mutate(expected_rides = 
           predict(rides_4kph, newdata = pred_df))

pred_df
```

The predicted ridership if the wind was blowing at 20 kph is 3515.342 rides.

```{r}
pred_df <- tribble(         # tribble allows easy manual data entry
  ~windspeed,
  20
)

pred_df <- pred_df |>
  mutate(expected_rides = predict(rides_4mph, newdata = pred_df))

pred_df
```

The predicted ridership if the wind was blowing at 20 mph is 3094.611 rides.

## Question 5

```{r}
bikes <- bikes %>%
  mutate(temp_c=((temp_feel-32)*(5/9)))
```

```{r}
newridesfit <- lm(rides ~ wind_kph+temp_c, data = bikes)

broom::tidy(newridesfit)
```

For a 1 kph increase in wind-speed, the model expects a 19.84 decrease in rides.

For a 1 degree increase in temperature, the model expects a 143.23 increase in rides.

## Question 6

```{r}
pred_df <- tibble(
  situation = 1:3,
  temp_c = c(25, 15, 10),
  wind_kph = c(15, 5, 40)
)

pred_df <- pred_df |>
  mutate(expected_rides = predict(newridesfit, newdata = pred_df))

pred_df
```

## Question 7

```{r}
bikes <- bikes |>
  mutate(weekend = as.factor(weekend))
```

```{r}
model2 <- lm(rides ~ temp_c + wind_kph + weekend, data = bikes)

get_regression_table(model2) |> select(term,estimate)
```

The coefficient "weekend" with the word TRUE next to it means that if its the weekend, then there will be -713.575 rides or 713.575 less rides on the weekend.

## Question 8

```{r}
pred_df <- tibble(
  situation = 1:2,
  temp_c = c(mean(bikes$temp_c), 
             mean(bikes$temp_c)),
  wind_kph = c(mean(bikes$wind_kph), 
               mean(bikes$wind_kph)), 
  weekend = c("TRUE","FALSE")
)

pred_df <- pred_df |>
  mutate(expected_rides = predict(model2, newdata = pred_df))

pred_df
```

This tells us that on a weekday (weekend is FALSE) there are 3683.442 expected rides.

## Question 9

```{r}
result <- get_regression_points(model2) %>%
  arrange(desc(abs(residual)))

result
```

The largest absolute residual was -3490.348. After doing some research, I found out that on this day was Hurricane Sandy which occurred in October 29, 2012.
