---
title: "HW MD 7"
format: html
editor: visual
embed-resources: true
---

## Homework MD

```{r}
library(tidyverse)
library(here)
library(moderndive)
theme_set(theme_minimal())


nobel_winners <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/nobel_winners.csv")
```

```{r}
nobel_winners_flt <- nobel_winners %>% 
  filter(category != "Peace") %>% 
  mutate(is_us = if_else(birth_country == "United States of America", 1, 0))
```

```{r}
true_prop <- nobel_winners_flt %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(nobel_winners_flt))

ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  labs(x = "Proportion")

```

```{r}
set.seed(33)
# Random sample 
our_sample <- sample_n(nobel_winners_flt, 
                       size = 25, 
                       replace = F)

sample_prop <- our_sample %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(our_sample))

# How does our proportion look like? 
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  geom_vline(xintercept = sample_prop[2,2][[1]], linetype = "dotted", col = "red") +
  labs(x = "Proportion")

```

```{r}
sample_prop
```

## Question 1

-   Nobel Prize Winners years 1901 - 2016.

<!-- -->

-   Nobel Prize Winners

<!-- -->

-   The black line represents the overall proportion of U.S Non-Peace Nobel Prize winners out of all of the Nobel Prize winners.

-   The red line represents the proportion of Non-Peace Nobel Prize winners that were born in the US out of a *sample* of 25 Non-Peace Nobel Prize winners.

## Question 2

```{r}
fifty_propershovel <- 
  nobel_winners_flt|> 
  rep_sample_n(size = 25,
               reps = 50) |>
  group_by(replicate) |>
  summarise(prop = sum(is_us)/25)
```

```{r}
ggplot(data = fifty_propershovel, 
       mapping = aes(x = prop))+
  geom_histogram(binwidth = 0.02, color = "navy")+
  geom_vline(aes(xintercept = mean(prop)),
             color="grey", linetype="dashed", size=.5) +
  labs(x = "Proportion")
```

## Question 3

```{r}
hunnid_propershovel <- 
  nobel_winners_flt|> 
  rep_sample_n(size = 100,
               reps = 50) |>
  group_by(replicate) |>
  summarise(prop = sum(is_us)/100)
```

```{r}
ggplot(data = hunnid_propershovel, 
       mapping = aes(x = prop))+
  geom_histogram(binwidth = 0.02, color = "navy")+
  geom_vline(aes(xintercept = mean(prop)),
             color="grey", linetype="dashed", size=.5) +
  labs(x = "Proportion")
```

There appears to be more variety in the distribution with 50 samples of 25 winners each over the distribution with 50 samples and 100 winners each. I think this is because as sample size increases, the variety of sampling distribution decreases, and we get more values closer to the average.

## Question 4

```{r}
std25 <- fifty_propershovel$prop      

sd(std25) 
```

```{r}
std100 <- hunnid_propershovel$prop      

sd(std100) 

```

The standard deviation tells above proves that the larger the sample size the lower the variation and therefore the lower the standard deviation.

## Question 5

```{r}
nobelwinners_women<- nobel_winners_flt |>
  mutate(gender1 = if_else(gender=="Female",1,0))
```

```{r}
female_100 <- nobelwinners_women |> 
  rep_sample_n(size = 25,
               reps = 50) |>
  group_by(replicate) |>
  summarise(prop = sum(gender1)/100)

ggplot(data =female_100, 
       mapping = aes(x = prop)) +
  geom_histogram(binwidth = 0.02, color = "navy") +
  geom_vline(aes(xintercept = mean(prop)),
             color="grey", linetype="dashed", size=.5) +
  labs(x = "Proportion")
```

```{r}
female_100 <- nobelwinners_women |> 
  rep_sample_n(size = 100,
               reps = 50) |>
  group_by(replicate) |>
  summarise(prop = sum(gender1)/100)

ggplot(data =female_100, 
       mapping = aes(x = prop)) +
  geom_histogram(binwidth = 0.02, color = "navy") +
  geom_vline(aes(xintercept = mean(prop)),
             color="grey", linetype="dashed", size=.5) +
  labs(x = "Proportion")
```

The proportion of women who have won the Nobel Prize is about 0.038-0.039 out of a sample of 100 scientists.

## Question 6

The central limit theorem is a theorem that states that as a sample size increases, the distribution of averages will appear more 'normal' and narrower. This will represent more of a bell-like function, and the higher sample size will more closely represent the true mean of a population's whole.
